---
title: "Filter_by_offshore_distance"
output:
  html_document:
      toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

We want to subset a dataset based on the distance from the coast. All near-shore points need to be excluded from the datasets, and based on this we want to explore the datasets.


## Create filter polygon

### Belgian part of the North Sea polygon

load libraries

```{r mregions2, echo=FALSE}
remotes::install_github('lifewatch/mregions2')
library(mapview)
library(mregions2)
library(sf)
library(dplyr)
library(tidyr)
```
Get the Belgian Part of the North Sea (also called 'Belgisch continentaal plat' or 'BCP')
```{r bcp}
BCP <- gaz_search(3293) %>% gaz_geometry()
mapview(BCP)
```
### Coastal area polygon

We will use a buffer around the baseline to remove the 'coastal' areas.

Get the baseline of Belgium
```{r download_baseline, echo=FALSE}
# Data from Agency Maritime Services and Coast (MDK)
# https://www.agentschapmdk.be/nl/publicaties#maritieme-grenzen-belgie

temp <- tempfile()
download.file("https://www.agentschapmdk.be/nl/bijlage/ec7480f9-7bdf-42a9-bcc1-d58f5a2a280c/maritieme-grenzen-be-december2018-4",temp)
unzip(temp, exdir = 'baseline')
bline <- st_read("baseline/SHAPE_basislijn BE -december2018.shp")
mapview(bline)
unlink(temp)

```

Now create a 1.5 km buffer around the baseline & union buffer with BCP.
You can change the buffer distance here, to have another distance from the coast.
```{r buffer_intersect}
buffer <- st_buffer(x = st_transform(bline, crs = 3812), # buffer, but transform to Lambert2008 (EPSG:3812) for distance
                    dist = 1500 ) # buffer distance in m

mapview(buffer)

BCP_1500 <- st_difference(BCP, st_transform(buffer, crs = 4326)) # transform to WGS84 (EPSG)
mapview(BCP_1500)

```
### Clean up polygon

There are still a remnant polygons in the harbour of Zeebrugge (and the harbour of Oostende), remove these.
```{r cleanp_poly}
#
BCP_1500_split <- BCP_1500 %>%
  st_cast(., "POLYGON") %>%             # convert multipolygon to seperate polygons
  mutate(area = st_area(the_geom)) %>%  # calculate area of separate polygons
  top_n(1, area)                        # only select largest polygon

mapview(BCP_1500_split)
```

## Dataset

### load data

Load the data (assumed to be in the project/working directory).
Note that this is a subset, as the full dataset (>700MB) is too large to load?
Note that I also subset the data to only include 2012.
```{r}

data <- read.csv("shiptemp_subset.csv",
                 sep = ";",
                 na.strings = "-999.0",
                 colClasses = c("POSIXct", # GPSDate to POSIXct format
                                "numeric",
                                "numeric",
                                "numeric",
                                "numeric",
                                "numeric",
                                "numeric")
                 )
data_sf <- data %>%
  drop_na('longitude') %>% # drop rows where lon is empty
  drop_na('Latitude') %>%  # drop rows where lat is empty
  mutate(year = format(GPSDate, format="%Y")) %>% # add year column
  filter(year == "2012") %>%                    # SELECT ONLY DATA IN 2012
  st_as_sf(coords = c("longitude", "Latitude"), # convert to sf object
           crs = 4326)
```

## Subset data
Subset the data based on the polygon
```{r}
data_BCP_1500 <- st_filter(data_sf, BCP_1500_split)
plot(st_geometry(BCP_1500_split))
plot(st_geometry(data_BCP_1500), col = 'red', add = TRUE)
```
(note that I don't load the data in a mapview, as this would make the html markdown output quite heavy)

## create some plots

Create some plots similar to Jonas' plots
```{r}
library(ggplot2)
ggplot(data_BCP_1500, aes(GPSDate, SBE21Temperature)) +
  geom_point(colour = "red",
             size = 0.2) +
  geom_hline(yintercept = 21,
             linetype = "dashed") +
  facet_wrap(vars(year),
             scales = "free_x") +
  theme_minimal()
```

